---
title: "Smart devices usage"
output: html_notebook
---

# CONTEXT
Bellabeat is a high-tech company that manufactures health-focused smart products since 2013. Their products:
* **Bellabeat app**: provides users with health data related to their activity, sleep, stress,
menstrual cycle, and mindfulness habits. It connects to their line of smart wellness products.
* **Leaf**: a wellness tracker can be worn as a bracelet, necklace, or clip. It connects to the Bellabeat app to track activity, sleep, and stress.
* **Time**: a wellness watch to track user activity, sleep, and stress. It connects to the Bellabeat app to provide daily wellness.
* **Spring**: a water bottle that tracks daily water intake that connects to the Bellabeat app to track
hydration levels.
* **Bellabeat membership**: a subscription-based membership program for users that gives 24/7 access to fully personalized guidance on nutrition, activity, sleep, health and beauty, and mindfulness based on their lifestyle and goals.

Bellabeat wants to get insights about the smart devices usage in order to obtain trends that help their marketing strategy.

# ASK

##Business task
Analyze smart device usage to gain insights into how people are already using their smart devices and provide high-level recommendations to improve marketing strategies for Bellabeat.

##Business use case
* Determine trends for smart device usage
* How could these trends apply to Bellabeat 'Time' customers?
* How could these trends help to improve Bellabeat's marketing strategy?

# PREPARE
For the analysis, I will use a custum survey (First-party data) in order to get updated data related to smartwatch usages, and use R with RStudio to perform the analysis task.

## Building the survey
### Platform
Build in Google Forms and available in 2 languages: English and Spanish. 
### Distribution
Distribuited trough a link using LinkedIn post, Instagram stories and Whatsapp groups. No restrictions to share the link.
### Time frame
Available for 4 days, starting on 01/26/2024
### Configuration
Anonymous survey
### Questions
#### Demographic questions
* Gender
* Age
* Geopraphic location

#### Smartwatch usage questions
* Main usage: multiple choice, with the option to add a custom answer
* Usage periodicity: scale from 0 to 7 representing usage in days
* Brand: multiple choice, listing main brand collected from sites like Amazon, Tiendamia, Mercado Libre. Option for add custom answer.
* Features used: checkboxs, listing main features offered by smartwatches on the market. Option for add custom answer.
* Tracked functionalities: checkboxs, listing main tracker options offers the smartwatches on the market. Option for add custom answer.
* Activities tracked: checkboxs, listing main sport activities tracked for the smartwatches on the market. Option for add custom answer.





## Evaluate data
Installing packages
```{r Installing packages, echo=FALSE}
install.packages("tidyverse")
library(tidyverse)

install.packages("skimr")
library(skimr)
```

Importing data
```{r Importing fitbit files}

```

Running basics evaluations over data
```{r Get a summary of daily_activity, echo=FALSE}
skim_without_charts(daily_activity)
str(daily_activity)
head(daily_activity)
```

# PROCESS
## Preparing environment for cleaning data
```{r Installing janitor package}
install.packages("janitor")
library(janitor)
```

Cleaning dataset names (names to snake_case)
```{r}
daily_activity_cleaned <- daily_activity %>% clean_names() %>% remove_empty(c("rows", "cols"))
```
daily_calories_cleaned <- clean_names(daily_calories)
daily_intensities_cleaned <- clean_names(daily_intensities)
daily_steps_cleaned <- clean_names(daily_steps)
daily_sleep_cleaned <- clean_names(daily_sleep)
seconds_heart_rate_cleaned <- clean_names(seconds_heart_rate)
weight_log_info_cleaned <- clean_names(weight_log_info)
```

